Structural controllability of ecological networks

### 2018-06-26 paper/reviewers-response.Rmd

> "Better context needs to be given for previous work to look at the management of networks in ecology. There are a few papers and knowing how this one differs from these (not just the ones that look at node importance for example) is important. Some of these papers are mentioned in the discussion but in the intro, it is important to know what they do and why this paper is novel."

Thanks for this observation. 
In line with feedback from both editors, we have chosen to emphasise on the conceptual aspects of the work and reduced the core question of management applications to appear only in the discussion. 
Nevertheless, we have also rewritten the introduction and took care to be clear at explaining how this study relates to and differs from previous work. 
In particular, we set up the stage for our work by briefly explaining how keystoneness metrics are mainly based on centrality metrics and how the proposed work differs from it.

> "Finally I think it is important to provide context about what kind of decisions are being made and could be informed by this approach. Who is managing plant-pollinator networks and what choices would they actually be making. Some examples should be expanded on. And also how would the outputs of this work actual inform choices and guide prioritization"

As it was explained in the previous response, in this paper we chose not to emphasize the empirical application of structural controllability. 
We acknowledge however that the approach has potential to be applied in management and have hence described some insights that structural controllability could offer to the conservation of ecological networks (\lr{M-management-implications}).

@@

> "And finally I think mentioning the extreme uncertainty mostly seen in networks in the real world is key, how can approaches deal with this needs to be discussed. What do we need to think about in terms of key information to use this approach?"

In the previous manuscript, we discussed this uncertainty but it was buried in the Supporting Information. 
We have now added a subsection in the Methods describing how we evaluate the impact of the uncertainty induced by using visitation as a proxy of species interdependence and that induced by incomplete sampling of rare interactions (\lr{M-testing-assumptions-line}). 
We found that the structural controllability approach is robust to these two sources of uncertainty (\lr{M-testing-assumptions-results-line}). 
Further details can be found in the Supporting Information Section \@ref(S-visitation-as-proxy) and \@ref(S-undersampling). 
In addition, we also included a brief description of the key information needed to apply the method in the Discussion (\lr{M-data-requirements}).

@@

As mentioned earlier in a response to the editor, we found that in order to address effectively all concerns, a rewrite of the whole paper was needed. 
Although we have not gone in the applied ecology direction, we hope the current version addresses the fundamental reviewers concerns satisfactorily.

> "One last minor remark, the authors have overlooked some key references in the field of restoration/conservation and ecological interactions/networks, which would strengthen the manuscript further, e.g.

> * Harvey, E., Gounand, I., Ward, C.L. & Altermatt, F. (2017) Bridging ecology and conservation: from ecological networks to ecosystem function. Journal of Applied Ecology, 54, 371-379; 
> * Devoto, M., Bailey, S., Craze, P. & Memmott, J. (2012) Understanding and planning ecological restoration of plant-pollinator networks. Ecology Letters, 15, 319-32; 
> * Kaiser-Bunbury, C.N., Mougal, J., Whittington, A.E., Valentin, T., Gabriel, R., Olesen, J.M. & Blüthgen, N. (2017) Ecosystem restoration strengthens pollination network resilience and function. Nature, 542, 223-227;
> * Pires, M.M., Marquitti, F.M.D. & Guimarães Jr, P.R. (2017) The friendship paradox in species-rich ecological networks: Implications for conservation and monitoring. Biological Conservation, 209, 245-252.

Thanks for these references. 
They have been particularly useful to highlight the differences between structural controllability and other approaches (\lr{M-extra-references-1} and \lr{M-extra-references-2}) and to discuss the potential impact of invasive plants (\lr{M-extra-references-3}).

### 2018-06-26 paper/manuscript.Rmd
Add references for line numbers
Close #40

Other conceptual differences between structural controllability and centrality metrics might provide further insight into the conservation of ecological networks. 
First, structural controllability emphasizes that the effect a species has on the abundance of other species is not independent of the effects of other other species in their community. 
Second, it demonstrates that to ensure the persistence of a community it is often necessary to consider the abundances of more than a single species, even when full control is unfeasible or undesired (for example `r round(sum(crit$n_crit > 1)/nrow(crit)*100)`% of our communities contained more than one critical species). 
Third, structural controllability explicitly acknowledges the existence of multiple management strategies and some will be better than others depending on the context.
Approaches to prioritise species for conservation and reintroduction based on traits or centrality are still useful and are likely to overlap with a species' control capacity [@pires_friendship_2017; @devoto_understanding_2012]. 
Altogether, our results also provide support to the idea that management decisions should not be based on a single technique, but indicates that focusing on ecosystem processes and interactions may be more effective than traditional ranking based approaches [@harvey_bridging_2017].

The choice of the invaded/uninvaded networks was based on our desire to contrast the extensive empirical evidence of the role of invasive plants with our theoretical results. 
We found that without exception, invasive plants were critical for network control and as such our results were in line with our expectations. 
Invasive plants have been previously found to exacerbate the asymmetries in their communities [@aizen_invasive_2008; bartomeus_contrasting_2008-1; @henriksson_strong_2016] and to be central in their communities [@vila_invasive_2009; @palacio_generalist_2016]. 
We found that invasive plants are, however, unlikely to be inherently different from their native counterparts [@emer_species_2016; @stouffer_how_2014]. 
Just like any other mutualist in our dataset, invasive plants tended to attain a high control capacity proportional to the degree to which they contribute to network persistence, are abundant, and depend little on other species. 
Furthermore, our observation that changes in the abundance of invasive plants (and presumably all critical species) are crucial to modify the state of the community agrees with recent evidence showing that ecosystem restoration focused on the eradication of invasive plants can have transformative desirable effects in plant-pollinator communities [@kaiser-bunbury_ecosystem_2017]. 
However, our results also suggest that removals must be exercised with caution. 
Not only it is hard to predict the direction in which the system will change, but we also show that critical species can underpin the coexistence of species and therefore some communities may be acutely vulnerable to their eradication [@traveset_invaders_2013; @albrecht_consequences_2014].

Structural stability assumes that the networks can be approximated using linear functional responses. 
The ramifications of this assumption imply that while structural controllability is useful to identify species that are relevant for network control, it cannot be used to design the precise interventions that should be applied to these species in order to achieve a desired state. 
While in an ideal scenario we would completely incorporate the species dynamics into the controllability analysis [@cornelius_realistic_2013], the reality is that such information is rarely available in most ecological scenarios. 
In contrast, structural controllability only requires a quantitative approximation of the network's interactions to gain valuable insight from the community.
Furthermore, unlike centrality metrics, the assumptions of structural controllability are explicit and the estimation of a species importance arises from a mechanistic understanding of the population dynamics between species. 
By accounting for network dynamics (even if in a simple way) it incorporates more ecological realism even in the extreme scenario where the state of a community depends almost exclusively on nodal dynamics rather than on the topology of their interactions.

>>>>

\R{management-implications}Other conceptual differences between structural controllability and centrality metrics might provide further insight into the conservation of ecological networks. 
First, structural controllability emphasizes that the effect a species has on the abundance of other species is not independent of the effects of other other species in their community. 
Second, it demonstrates that to ensure the persistence of a community it is often necessary to consider the abundances of more than a single species, even when full control is unfeasible or undesired (for example `r round(sum(crit$n_crit > 1)/nrow(crit)*100)`% of our communities contained more than one critical species). 
Third, structural controllability explicitly acknowledges the existence of multiple management strategies and some will be better than others depending on the context.
Approaches to prioritise species for conservation and reintroduction based on traits or centrality are still useful and are likely to overlap with a species' control capacity\R{extra-references-1} [@pires_friendship_2017; @devoto_understanding_2012]. 
Altogether, our results also provide support to the idea that management decisions should not be based on a single technique, but indicates that focusing on ecosystem processes and interactions may be more effective than traditional ranking based approaches\R{extra-references-2} [@harvey_bridging_2017].

The choice of the invaded/uninvaded networks was based on our desire to contrast the extensive empirical evidence of the role of invasive plants with our theoretical results. 
We found that without exception, invasive plants were critical for network control and as such our results were in line with our expectations. 
Invasive plants have been previously found to exacerbate the asymmetries in their communities [@aizen_invasive_2008; bartomeus_contrasting_2008-1; @henriksson_strong_2016] and to be central in their communities [@vila_invasive_2009; @palacio_generalist_2016]. 
We found that invasive plants are, however, unlikely to be inherently different from their native counterparts [@emer_species_2016; @stouffer_how_2014]. 
Just like any other mutualist in our dataset, invasive plants tended to attain a high control capacity proportional to the degree to which they contribute to network persistence, are abundant, and depend little on other species. 
Furthermore, our observation that changes in the abundance of invasive plants (and presumably all critical species) are crucial to modify the state of the community agrees with recent evidence showing that ecosystem restoration focused on the eradication of invasive plants can have transformative desirable effects in plant-pollinator communities\R{extra-references-3} [@kaiser-bunbury_ecosystem_2017]. 
However, our results also suggest that removals must be exercised with caution. 
Not only it is hard to predict the direction in which the system will change, but we also show that critical species can underpin the coexistence of species and therefore some communities may be acutely vulnerable to their eradication [@traveset_invaders_2013; @albrecht_consequences_2014].

Structural stability assumes that the networks can be approximated using linear functional responses. 
The ramifications of this assumption imply that while structural controllability is useful to identify species that are relevant for network control, it cannot be used to design the precise interventions that should be applied to these species in order to achieve a desired state. 
While in an ideal scenario we would completely incorporate the species dynamics into the controllability analysis [@cornelius_realistic_2013], the reality is that such information is rarely available in most ecological scenarios. 
\R{data-requirements}In contrast, structural controllability only requires a quantitative approximation of the network's interactions to gain valuable insight from the community.
Furthermore, unlike centrality metrics, the assumptions of structural controllability are explicit and the estimation of a species importance arises from a mechanistic understanding of the population dynamics between species. 
By accounting for network dynamics (even if in a simple way) it incorporates more ecological realism even in the extreme scenario where the state of a community depends almost exclusively on nodal dynamics rather than on the topology of their interactions.
